No need to know: astrophysics-agnostic gravitational wave cosmology

### 2024-03-11 src/tex/ms.tex
[showyourwork] overleaf sync

The mass distribution inferred with the \ac{GP}-based model closely resembles the true, simulated distribution.%, and the injected value of \Ho{} ($\variable{output/H0_FID.txt}\Hunits$) is consistent with its inferred posterior ($\variable{output/nonparh0CI.txt}\Hunits$).
At very high and very low masses, there is both a lack of data and little detector sensitivity.
The mass distribution is therefore uninformed in this region and the \ac{GP} reverts to its prior distribution, which resembles random scatter around the mean differential merger rate.
Similar effects can be seen in other non-parametric methods \citep{edelman_cover_2023, callister_parameter-free_2023}.
On the other hand, the lack of data just below the minimum \ac{BH} mass and just above the maximum \ac{BH} mass, combined with the fact that \ac{GW} detectors are sensitive to objects in that mass range causes the \ac{GP} to learn a relatively low merger rate at the edges of the mass distribution.
Indeed, these features are very informative to the \Ho{} constraint \citep{the_ligo_scientific_collaboration_constraints_2021}.
The combination of these two effects results in what appears as an uptick in the merger rate at very low and very high masses, but these features are uninformative and do not affect inference on cosmological parameters.
Additionally, the posterior on \Ho{} is distinct from its prior distribution (uniform in the range $[30\Hunits,60\Hunits]$, indicating that the data is informative despite the flexibility of the population model.

@@

The resulting inferred mass distribution is shown in Figure~\ref{fig:O5_GP}, along with the corresponding posterior on \Ho.
The inferred mass distribution closely resembles the true, simulated distribution, and the injected value of \Ho{} ($\variable{output/H0_FID.txt}\Hunits$) is consistent with its inferred posterior ($\variable{output/nonparh0CI.txt}\Hunits$).
At very high and very low masses, there is both a lack of data and little detector sensitivity.
The mass distribution is therefore uninformed in this region and the \ac{GP} reverts to its prior distribution, which resembles random scatter around the mean differential merger rate.
Similar effects can be seen in other non-parametric methods \citep{edelman_cover_2023, callister_parameter-free_2023}.
On the other hand, the lack of data just below the minimum \ac{BH} mass and just above the maximum \ac{BH} mass, combined with the fact that \ac{GW} detectors are sensitive to objects in that mass range causes the GP to learn a relatively low merger rate at the edges of the mass distribution.
Indeed, these features are very informative to the \Ho{} constraint \citep{the_ligo_scientific_collaboration_constraints_2021}.
The combination of these two effects results in what appears as an uptick in the merger rate at very low and very high masses, but these features are uninformative and do not affect inference on cosmological parameters.
Additionally, the posterior on \Ho{} is distinct from its prior distribution (uniform in the range $[30\Hunits,60\Hunits]$, indicating that the data is informative despite the flexibility of the population model.

>>>>

The mass distribution inferred with the \ac{GP}-based model closely resembles the true, simulated distribution.%, and the injected value of \Ho{} ($\variable{output/H0_FID.txt}\Hunits$) is consistent with its inferred posterior ($\variable{output/nonparh0CI.txt}\Hunits$).
At very high and very low masses, there is both a lack of data and little detector sensitivity.
The mass distribution is therefore uninformed in this region and the \ac{GP} reverts to its prior distribution, which resembles random scatter around the mean differential merger rate.
Similar effects can be seen in other non-parametric methods \citep{edelman_cover_2023, callister_parameter-free_2023}.
On the other hand, the lack of data just below the minimum black hole mass and just above the maximum black hole mass, combined with the fact that \ac{GW} detectors are sensitive to objects in that mass range causes the \ac{GP} to learn a relatively low merger rate at the edges of the mass distribution.
Indeed, these features are very informative to the \Ho{} constraint \citep{the_ligo_scientific_collaboration_constraints_2021}.
The combination of these two effects results in what appears as an uptick in the merger rate at very low and very high masses, but these features are uninformative and do not affect inference on cosmological parameters.
Additionally, the posterior on \Ho{} is distinct from its prior distribution (uniform in the range $[30\Hunits,60\Hunits]$, indicating that the data is informative despite the flexibility of the population model.

@@

The resulting inferred mass distribution is shown in Figure~\ref{fig:O5_GP}, along with the corresponding posterior on \Ho.
The inferred mass distribution closely resembles the true, simulated distribution, and the injected value of \Ho{} ($\variable{output/H0_FID.txt}\Hunits$) is consistent with its inferred posterior ($\variable{output/nonparh0CI.txt}\Hunits$).
At very high and very low masses, there is both a lack of data and little detector sensitivity.
The mass distribution is therefore uninformed in this region and the \ac{GP} reverts to its prior distribution, which resembles random scatter around the mean differential merger rate.
Similar effects can be seen in other non-parametric methods \citep{edelman_cover_2023, callister_parameter-free_2023}.
On the other hand, the lack of data just below the minimum black hole mass and just above the maximum black hole mass, combined with the fact that \ac{GW} detectors are sensitive to objects in that mass range causes the GP to learn a relatively low merger rate at the edges of the mass distribution.
Indeed, these features are very informative to the \Ho{} constraint \citep{the_ligo_scientific_collaboration_constraints_2021}.
The combination of these two effects results in what appears as an uptick in the merger rate at very low and very high masses, but these features are uninformative and do not affect inference on cosmological parameters.
Additionally, the posterior on \Ho{} is distinct from its prior distribution (uniform in the range $[30\Hunits,60\Hunits]$, indicating that the data is informative despite the flexibility of the population model.
